---
title: "Assess stage-discharge data"
author: "Maartje Korver"
date: "May 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

# Explanation of code
## Action required

# Install and/or load packages
#install.packages('ggplot2')
#install.packages('plotly')
#install.packages('tidyverse')
#install.packages('googlesheets')

library(ggplot2)
library(plotly)
library(tidyverse)
library(googlesheets)

## Set your working directory
setwd("C:/Users/Watersheds/Documents/RProjects/Rating_curves/Test")

```

```{r load_data, include=FALSE}

## Load stage-discharge data from googlesheet. Adapt sheet title (gs_title) and tab name (gs_read).
HQ <- gs_title('Metadata Rating curve 708')
HQ <- gs_read(HQ, ws = "Rating curve v4")

# Add column with absolute Q uncertainty. 
# Filter measurements that are indicated as 'Y' under 'Final_rating_curve'
HQ <- HQ %>% 
  mutate(Q_abs_unc = Q_rel_unc/100 * Q_meas) %>% 
  filter(Final_rating_curve == 'Y')

## Load the rating curve lookup table of the previous rating curve version. Adapt sheet title (gs_title) and tab name (gs_read).
ratingcurve_old <- gs_title('Rating curves v3 - H-Q lookup tables')
ratingcurve_old <- gs_read(ratingcurve_old, ws = "WTS708")

# Create separate dataframes for different rating shifts
ratingcurve_old_rc1 <- ratingcurve_old %>%
  filter(Rating == '1')
ratingcurve_old_rc2 <- ratingcurve_old %>%
  filter(Rating == '2')
#ratingcurve_old_rc2 <- ratingcurve_old %>%
  #filter(Rating == '3')

# Create dataframes filtered for old or new storm events
HQ_newstorm <- HQ %>% 
  filter(Old_new == 'New')
HQ_oldstorm <- HQ %>% 
  filter(Old_new == 'Old')

# Create dataframes calculating residuals of Q_meas and Q_model
resid_new <- HQ_newstorm %>% 
  mutate(Stage_avg = round(Stage_avg, digits = 1)) %>% 
  select(Stage_avg, Q_meas)
resid_old <- HQ_oldstorm %>% 
  mutate(Stage_avg = round(Stage_avg, digits = 1)) %>% 
  select(Stage_avg, Q_meas)

resid_new <- merge(resid_new, ratingcurve_old_rc2, by.x = 'Stage_avg')
resid_old <- merge(resid_old, ratingcurve_old_rc2, by.x = 'Stage_avg')

resid_new <- resid_new %>% 
  mutate(resid_new = Q_meas - Q_model)
resid_old <- resid_old %>% 
  mutate(resid_old = Q_meas - Q_model)

```

### Watershed 708 -- version 4

_The rating curve plots in this document are not final, but used as a tool for measurement selection and identification of rating curve shifts. For detailed measurement information see [Metadata Rating Curve 708](https://docs.google.com/spreadsheets/d/1yCLgQHZ0vxCkPNo4wxwIYnYjdTSFAqXOZcao8AgZQZ8/edit#gid=1712786160)_

* `r nrow(HQ_newstorm)` new discharge measurements were taken since the last rating curve update, from `r (max(HQ_newstorm$Event_no)-min(HQ_newstorm$Event_no)+1)` different storm events.
* Total number of discharge measurements = `r nrow(HQ)`


#### Overview

```{r new_vs_old, include=TRUE}

ggplot() + 
  geom_line(data = ratingcurve_old_rc1, aes(x = Stage_avg, y = Q_model)) + 
  geom_line(data = ratingcurve_old_rc2, aes(x = Stage_avg, y = Q_model)) +
  #geom_line(data = ratingcurve_old_rc3, aes(x = Stage_avg, y = Q_model)) +
  geom_point(data = HQ, aes(x = Stage_avg, y = Q_meas, fill = Old_new), shape = 21, size = 2) + 
  theme_bw() + scale_x_continuous(minor_breaks = seq(0, 200, 10)) + facet_grid(Old_new ~ .) + guides(fill = "none") + xlab("Stage [cm]") + ylab("Discharge [m3/s]")

```

#### Zoom plots {.tabset}

##### New vs. old
```{r new_vs_old_2, include=TRUE}

ggplotly(ggplot() + 
  geom_line(data = ratingcurve_old_rc1, aes(x = Stage_avg, y = Q_model)) + 
  geom_line(data = ratingcurve_old_rc2, aes(x = Stage_avg, y = Q_model)) + 
  #geom_line(data = ratingcurve_old_rc3, aes(x = Stage_avg, y = Q_model)) + 
  geom_pointrange(data = HQ, aes(x = Stage_avg, y = Q_meas, ymin = Q_meas - Q_abs_unc, ymax = Q_meas + Q_abs_unc, color = Old_new, Label = ID), size = 1.5) + 
  theme_bw() + xlab("Stage [cm]") + ylab("Discharge [m3/s]"))

```

##### Storm events
```{r new_storm, include=TRUE}

ggplotly(ggplot() + 
  geom_line(data = ratingcurve_old_rc1, aes(x = Stage_avg, y = Q_model)) + 
  geom_line(data = ratingcurve_old_rc2, aes(x = Stage_avg, y = Q_model)) +
  #geom_line(data = ratingcurve_old_rc3, aes(x = Stage_avg, y = Q_model)) +
  geom_point(data = HQ_newstorm, aes(x = Stage_avg, y = Q_meas, fill = factor(Event_no), Label = ID), shape = 21, size = 2) + 
  theme_bw() + xlab("Stage [cm]") + ylab("Discharge [m3/s]") + labs(fill = 'Storm event no.'))

```

##### Water Year
```{r shifts, include=TRUE}

ggplotly(ggplot() + 
  geom_line(data = ratingcurve_old_rc1, aes(x = Stage_avg, y = Q_model)) + 
  geom_line(data = ratingcurve_old_rc2, aes(x = Stage_avg, y = Q_model)) +
  #geom_line(data = ratingcurve_old_rc3, aes(x = Stage_avg, y = Q_model)) +
  geom_point(data = HQ, aes(x = Stage_avg, y = Q_meas, fill = factor(WTY), Label = ID), shape = 21, size = 2) + 
  theme_bw() + xlab("Stage [cm]") + ylab("Discharge [m3/s]") + labs(fill = 'Shift segment'))

```

##### Hysteresis
```{r hysteresis, include=TRUE}

ggplotly(ggplot() + 
  geom_line(data = ratingcurve_old_rc1, aes(x = Stage_avg, y = Q_model)) + 
  geom_line(data = ratingcurve_old_rc2, aes(x = Stage_avg, y = Q_model)) +
  #geom_line(data = ratingcurve_old_rc3, aes(x = Stage_avg, y = Q_model)) +
  geom_point(data = HQ, aes(x = Stage_avg, y = Q_meas, fill = Stage_delta, Label = ID), shape = 21, size = 2) + 
  theme_bw() + xlab("Stage [cm]") + ylab("Discharge [m3/s]") + labs(fill = 'Falling or rising limb'))

```

#### Did Rating curve shift? {.tabset}

_To assess whether the rating curve has shifted (objectively), the residuals (Q_meas - Q_model) of the discharge measurements to the old rating curve model are assessed. If the mean of all residuals is larger then the mean of all absolute measurement uncertainties, a new shift in the rating curve should be considered._

##### Old
```{r resid_old, include = TRUE}
cols <- c("Q_resid" = "black", "mean Q_abs_unc" = "grey","mean Q_resid" = "red")

ggplot() + 
  geom_line(data = resid_old, aes(x = Stage_avg, y = resid_old, color = "Q_resid")) +
  geom_point(data = resid_old, aes(x = Stage_avg, y = resid_old, color = "Q_resid")) +
  geom_hline(aes(yintercept = mean(HQ_oldstorm$Q_abs_unc), color = "mean Q_abs_unc"), size = 1, show.legend = TRUE) +
  geom_hline(aes(yintercept = -mean(HQ_oldstorm$Q_abs_unc), color = "mean Q_abs_unc"), size = 1, show.legend = TRUE) + 
  geom_hline(aes(yintercept = mean(resid_old$resid_old), color = "mean Q_resid"), size = 1, show.legend = TRUE) + 
  theme_bw() + xlab("Stage [cm]") + ylab("residuals [m3/s]") + scale_color_manual(name = "", values = cols)
```

##### New
```{r resid_new, include = TRUE}
cols <- c("Q_resid" = "black", "mean Q_abs_unc" = "grey","mean Q_resid" = "red")

ggplot() + 
  geom_line(data = resid_new, aes(x = Stage_avg, y = resid_new, color = "Q_resid")) +
  geom_point(data = resid_new, aes(x = Stage_avg, y = resid_new, color = "Q_resid")) +
  geom_hline(aes(yintercept = mean(HQ_newstorm$Q_abs_unc), color = "mean Q_abs_unc"), size = 1, show.legend = TRUE) +
  geom_hline(aes(yintercept = -mean(HQ_newstorm$Q_abs_unc), color = "mean Q_abs_unc"), size = 1, show.legend = TRUE) + 
  geom_hline(aes(yintercept = mean(resid_new$resid_new), color = "mean Q_resid"), size = 1, show.legend = TRUE) + 
  theme_bw() + xlab("Stage [cm]") + ylab("residuals [m3/s]") + scale_color_manual(name = "", values = cols)
```

