---
title: "Final rating curve"
author: "Maartje Korver"
date: "May 13, 2019"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Explanation of code
## Action required

# Install and/or load packages
#install.packages('ggplot2')
#install.packages('plotly')
#install.packages('tidyverse')
#install.packages('stats')
#install.packages('googlesheets')
#install.packages('readr')
#install.packages('KScorrect')

library(ggplot2)
library(plotly)
library(tidyverse)
library(stats)
library(googlesheets)
library(readr)
library(KScorrect)

## Set working directory
setwd("C:/Users/Watersheds/Documents/RProjects/Rating_curves/Test")

# Read functions
source("selectspan.R")
source("interp1.R")
source("interp2.R")
source("HQ_unc.R")
source("CI_model_input.R")
source("CI.R")

```

```{r load_data, include = FALSE}

## Load stage-discharge data from googlesheet. Adapt sheet title (gs_title) and tab name (gs_read).
HQ <- gs_title('Metadata Rating curve 626')
HQ <- gs_read(HQ, ws = "Rating curve v4")

# Add column with absolute Q uncertainty. 
# Filter measurements that are indicated as 'Y' under 'Final_rating_curve'
HQ <- HQ %>% 
  mutate(Q_abs_unc = Q_rel_unc/100 * Q_meas) %>% 
  filter(Final_rating_curve == 'Y')

```

### Watershed 626 -- version 4

_This document outlines the final rating curve plotting process, the decisions that were made and the results_

#### Conclusions drawn from Stage-discharge data assessment

**Were data points omitted?**  
No new measurements were omitted from this rating curve.
Two high flow measurements taken 2016-11-08 still plot far off the curve and are not included.

**Other observations/comments**  
Slight shift down can be observed in latest data. Not enough to plot a shifted curve.

#### Loess fit with visual span width selection

Select the span width that does not overfit (span width too small) nor underfit (span width too large) the data.

Span_a = 0.14  
Span_b = 0.15  
Span_c = 0.16  

Span width chosen = **span_b**

```{r define_span1, include = TRUE}

## Define potential span widths
Rating <- selectspan(HQ,0.14,0.15,0.16)

ggplotly(ggplot() +
  geom_pointrange(data = HQ, aes(x = Stage_avg, y = Q_meas, ymin = Q_meas - Q_abs_unc, ymax = Q_meas + Q_abs_unc, Label = ID), size = 1) +
  geom_line(data = Rating, aes(x = Stage_avg, y = Q_model, colour = smoothed)) + 
  theme_bw() + xlab("Stage [cm]") + ylab("Discharge [m3/s]") + labs(title = "Loess fit - visual span selection"))

```

```{r choose_span, include = FALSE}
## Enter chosen span width (span_a, span_b or span_c)
Rating <- Rating %>% 
  filter(smoothed == 'span_b') %>% 
  select(Stage_avg, Q_model)

## Enter span width value
Span <- 0.15

```

#### Extrapolation of curve to minimum and maximum stage {.tabset}

Extrapolation is based on the highest and lowest measured stage values to date:

| Stage [cm]| Date recorded |
|:----------|:--------------|
|   17.3    |   2018-08-24  |
|   204     |   2018-12-29  |

Shape/equation used for extrapolation: **power law equation**

*Additional information*:  
The extrapolation was based on a power law equation that was plotted on all measurements above an inflection point of 123 cm.
For details see **data_input/extrapolation_determination.xlsx** 

All extrapolated and estimated stage-discharge values can be found at   **data_input/extrap_values.csv**  

```{r extrap, include = TRUE}
## Estimate values for extrapolation (do separately fom this script) and write into 'extrap_values.csv' file.

## Import extrap_values from .csv file
extrap_values <- read_csv("data_input/extrap_values.csv")
extrap_values <- extrap_values %>%
  select(Stage_avg, Q_model)

# Add extrap_values to Rating
Rating_extrap <- Rating %>% 
  select(Stage_avg, Q_model)
Rating_extrap <- bind_rows(Rating_extrap, extrap_values)

# Plot Q measurements and modelled Q as well as the extrapolated part of modelled Q
ggplotly(ggplot() + 
  geom_line(data = Rating_extrap, aes(x = Stage_avg, y = Q_model), color = "red") + 
  geom_point(data = Rating_extrap, aes(x = Stage_avg, y = Q_model), color = "red") + 
  geom_point(data = HQ, aes(x = Stage_avg, y = Q_meas, Label = ID)) + 
  theme_bw() + xlab("Stage [cm]") + ylab("Discharge [m3/s]") + labs(title = "Estimated data points for extrapolation of curve"))

```

```{r interp1, include = FALSE}

# Interpolate modelled Q values (including extrapolated part) for each mm of stage
Rating_extrap_interp <- interp1(Rating_extrap)
names(Rating_extrap_interp) <- c("Stage_avg", "Q_model")

```

#### Confidence intervals

* Uncertainties of Q measurements are adapted by including stage uncertainty
* 500 datasets are created with randomly chosen Q_meas, Q + unc or Q - unc for each stage value
* 500 loess regressions are run using these randomized datasets
* 95% CI of 500 regression results are calculated
* Confidence intervals of the extrapolated part of the rating curve are estimated to be 5% of modelled discharge.

```{r CI, include = FALSE}

## Determine number of loess regressions for CI plotting - Coxon et al. (2015) suggests 500
n <- 500

# Calculate Q_meas uncertainty taking into account uncertainty in stage and create dataframe with Stage_avg, Q_meas and Q_H_abs_unc
HQ_unc <- HQ_unc(HQ, Rating, Rating_extrap_interp)

# Prepare n number of H-Q datasets, sampled from Q, Q_max, Q_min, H, H_max, H_min
CI_model_input <- CI_model_input(HQ_unc, n, HQ)

# Define Expand value (Expansion of qmixnorm() in CI.R). Default is 1. If CI.R produces NaN values, decrease value. Choose largest number that does not produce NaN values.
Expand <- 1

# Calculate 95% CI
CI <- CI(CI_model_input,n,Span,Expand)

# Create dataframe with Stage_avg, Q_model, Max_CI and Min_CI
RatingCI <- bind_cols(Rating, CI)

write_csv(RatingCI, "data_output/RatingCI.csv")

```

```{r extrap CI, include = FALSE}

## Add estimates for Max_CI and Min_CI of the extrapolated parts of the rating curve to the extrap_values.csv file (5% of modelled Q).
extrap_values <- read_csv("data_input/extrap_values.csv")

RatingCI_extrap <- bind_rows(RatingCI, extrap_values)

```

```{r interp2, include = FALSE}
# Interpolate values within measurement points
RatingCI_extrap_interp <- interp2(RatingCI_extrap)

```

#### Final rating curve

Results are written to: __data_output/WTS626ratingcurve_v4.csv__

```{r finish, include=TRUE}

## Adapt watershed name and version number and write data
write_csv(RatingCI_extrap_interp,"data_output/WTS626ratingcurve_v4.csv")
write_csv(HQ,"data_output/WTS626stagedischarge_v4.csv")

# Re-organizing dataframe for plotting
Q_meas <- cbind(HQ$Stage_avg,HQ_unc)
colnames(Q_meas)[1] <- "Stage_avg"

## Plot rating curve displaying CI and measurement uncertainties. Adapt plot title.
ggplotly(ggplot() + 
  geom_ribbon(data = RatingCI_extrap_interp, aes(x = Stage_avg, ymin = Min_CI, ymax = Max_CI), fill = "grey") +
  geom_pointrange(data = Q_meas, aes(x = Stage_avg, y = Q_meas, ymin = Q_min, ymax = Q_max), size = 1) +
  geom_line(data = RatingCI_extrap_interp, aes(x = Stage_avg, y = Q_model), colour = "#333333") +
  theme_bw() + xlab("Stage [cm]") + ylab("Discharge [m3/s]") + labs(title = "WTS626 - Rating curve"))

```
